---
title: "Integration Absolute Error"
author: 
  - Tsz Keung Wong
date: "2026-05-08"
output: 
  html_document:
    toc: true
    theme: united
    toc_float: true
editor_options: 
  chunk_output_type: console
---

This document evaluates the numerical accuracy of integration used in the `BayesPower` package. In particular, the focus is on the **absolute error** returned by `stats::integrate()` when computing expected power.


## Loading the required packages

```{r message=FALSE, warning=FALSE, results='hide'}
library("BayesPower")
```

## $t$-tests

The functions in the backend for the calculation of power, `BayesPower:::t1_TPE` and `BayesPower:::t2_TPE`, are modified such that they return the absolute error from the numerical integration instead of the estimated value of the integral.


```{r, results='hide'}
##### Helper #####

# one-sample t-test 
t1_TPE_abs_error <- function(t, df, prior_analysis, location, scale, dff, alternative ) {
  if (any(t == "bound cannot be found") || length(t) == 0) return(0)

  constant <- sqrt(df + 1)

  if (prior_analysis == "Point") {

    ncp <- location * constant

    pro <- switch(
      alternative,

      "two.sided" = {
        if (length(t) == 2) {
          BayesPower:::pnct(min(t), df, ncp = ncp, lower = TRUE) +
            BayesPower:::pnct(max(t), df, ncp = ncp, lower = FALSE)
        } else {
          if (t > 0) {
            BayesPower:::pnct(t, df, ncp = ncp, lower = FALSE)
          } else {
            BayesPower:::pnct(t, df, ncp = ncp, lower = TRUE)
          }
        }
      },

      "greater" = BayesPower:::pnct(t, df, ncp = ncp, lower = FALSE),

      "less" = BayesPower:::pnct(t, df, ncp = ncp, lower = TRUE)
    )

    return(pro)
  }
 # limits of the integral, parameter space
  bound <- switch(
    alternative,
    "greater" = c(a = 0, b = Inf),
    "less" = c(a = -Inf, b = 0),
    "two.sided" = c(a = -Inf, b = Inf)
  )
 # normalization of the prior under h1
  normalization <- if (alternative == "two.sided") {
    1
  } else {
    switch(
      prior_analysis,
      "Cauchy" = stats::pcauchy(bound[2], location, scale) -
        stats::pcauchy(bound[1], location, scale),

      "Normal" = stats::pnorm(bound[2], location, scale) -
        stats::pnorm(bound[1], location, scale),

      "Moment" = if (bound[2] == 0) {
        BayesPower:::pmom(bound[2] - location, tau = scale^2)
      } else {
        1 - BayesPower:::pmom(bound[1] - location, tau = scale^2)
      },

      "t-distribution" = stats::pt((bound[2] - location) / scale, dff, 0) -
        stats::pt((bound[1] - location) / scale, dff, 0)
    )
  }
 # integral, marginalized probability 
  int <- function(delta) {

    ncp <- delta * constant

    pro <- switch(
      alternative,

      "two.sided" = {
        if (length(t) == 2) {
          BayesPower:::pnct(min(t), df, ncp = ncp, lower = TRUE) +
            BayesPower:::pnct(max(t), df, ncp = ncp, lower = FALSE)
        } else {
          if (t > 0) {
            BayesPower:::pnct(t, df, ncp = ncp, lower = FALSE)
          } else {
            BayesPower:::pnct(t, df, ncp = ncp, lower = TRUE)
          }
        }
      },

      "greater" = BayesPower:::pnct(t, df, ncp = ncp, lower = FALSE),

      "less" = BayesPower:::pnct(t, df, ncp = ncp, lower = TRUE)
    )

    pro * BayesPower:::t1_prior(delta, location, scale, dff, prior_analysis) / normalization
  }

  error <- 1e-4

  if (prior_analysis == "Moment" && scale < 0.3) {
    error <- 1e-14
  }

  #stats::integrate(
  #  int,
  #  lower = bound[1],
  #  upper = bound[2],
  #  rel.tol = error,
  #  stop.on.error = FALSE
  #)$value
  stats::integrate(int, lower = bound[1], upper = bound[2], rel.tol = error, stop.on.error = FALSE)$abs.error
  
}

# Indenpedent samples t-test
t2_TPE_abs_error <-function(t,n1,r,prior_analysis ,location ,scale,dff , alternative ){
   n2 <- n1 * r
  df <- n1 + n2 - 2
  constant <- sqrt((n1 * n2) / (n1 + n2))

  if (any(t == "bound cannot be found") || length(t) == 0) return(0)

  if (prior_analysis == "Point") {

    pro <- switch(
      alternative,

      "two.sided" = {
        if (length(t) == 2) {
          BayesPower:::pnct(min(t), df, ncp = location * constant, lower = TRUE) +
            BayesPower:::pnct(max(t), df, ncp = location * constant, lower = FALSE)
        } else {
          if (t > 0 ) {
            BayesPower:::pnct(t, df, ncp = location * constant, lower = FALSE)
          } else {
            BayesPower:::pnct(t, df, ncp = location * constant, lower = TRUE)
          }
        }
      },

      "greater" = BayesPower:::pnct(t, df, ncp = location * constant, lower = FALSE),

      "less" = BayesPower:::pnct(t, df, ncp = location * constant, lower = TRUE)
    )

    return(pro)
  }

  # limits of the integral and the parameter space under h1
  bound <- switch(
    alternative,
    "greater" = c(a = 0, b = Inf),
    "less" = c(a = -Inf, b = 0),
    "two.sided" = c(a = -Inf, b = Inf)
  )

  # normalization of prior under h1
  normalization <- if (alternative == "two.sided") {
    1
  } else {
    switch(
      prior_analysis,
      "Cauchy" = stats::pcauchy(bound[2], location, scale) -
        stats::pcauchy(bound[1], location, scale),

      "Normal" = stats::pnorm(bound[2], location, scale) -
        stats::pnorm(bound[1], location, scale),

      "Moment" = if (bound[2] == 0) {
        BayesPower:::pmom(bound[2] - location, tau = scale^2)
      } else {
        1 - BayesPower:::pmom(bound[1] - location, tau = scale^2)
      },

      "t-distribution" = stats::pt((bound[2] - location) / scale, dff, 0) -
        stats::pt((bound[1] - location) / scale, dff, 0)
    )
  }

  # integral
  int <- function(delta) {

    ncp <- delta * constant

    pro <- switch(
      alternative,

      "two.sided" = {
        if (length(t) == 2) {
          BayesPower:::pnct(min(t), df, ncp = ncp, lower = TRUE) +
            BayesPower:::pnct(max(t), df, ncp = ncp, lower = FALSE)
        } else {
          if (t > 0) {
            BayesPower:::pnct(t, df, ncp = ncp, lower = FALSE)
          } else {
            BayesPower:::pnct(t, df, ncp = ncp, lower = TRUE)
          }
        }
      },

      "greater" = BayesPower:::pnct(t, df, ncp = ncp, lower = FALSE),

      "less" = BayesPower:::pnct(t, df, ncp = ncp, lower = TRUE)
    )

    pro * BayesPower:::t1_prior(delta, location, scale, dff, prior_analysis) / normalization
  }

  error <- 1e-4

  if (prior_analysis == "Moment" && scale < 0.3) {
    error <- 1e-14
  }

  #x <- stats::integrate(
  #  int,
  #  lower = bound[1],
  #  upper = bound[2],
  #  rel.tol = error,
  #  stop.on.error = FALSE
  #)$value
  
x <- stats::integrate(int,lower = bound[1],upper = bound[2], rel.tol = error,stop.on.error=FALSE)$abs.error
  return(x)

}
```

### Analysis settings

This section defines the prior specifications, sample-size settings, and critical values used in the numerical comparison. 

```{r, results='hide'}
df             <- 49
location       <- 0
scale          <- 1
dff            <- 1
t.cri          <- seq(0.1,5,.01)
prior_analysis <- c("t-distribution","Normal","Moment")
alternative    <- "two.sided"
n1             <- 50
r              <- 1
```

### Calculating absolute error across different $t$-values

```{r}
abs_error_t1 <- array(NA, dim=c(3,length(t.cri)))
abs_error_t2 <- array(NA, dim=c(3,length(t.cri)))

for( ii in 1:length(prior_analysis)){
for( i in 1:length(t.cri)){
  t <- c(-t.cri[i],t.cri[i]) # specifying which critical value is used for power calculation
  abs_error_t1[ii,i] <- t1_TPE_abs_error(t, df, prior_analysis[ii], location, scale, dff,alternative)
  abs_error_t2[ii,i] <- t2_TPE_abs_error(t,n1,r,prior_analysis[ii] ,location ,scale,dff , alternative )
}}
max(abs_error_t1,abs_error_t2)
```





## Correlation

The function in the backend for the calculation of power, `r_TPE`, is modified such that they return the absolute error from the numerical integration instead of the estimated value of the integral.


```{r, results='hide'}
##### Helper #####

r_TPE_abs_error <-function(r,n,k, alpha, beta,h0,alternative,location,scale,dff,prior_analysis){

    if (any(r == "bound cannot be found") || length(r) == 0) return(0)

  if (prior_analysis == "Point") {
    x <- switch(alternative,
                "two.sided" = if (length(r) == 2) BayesPower:::p_cor(max(r), location, n, lower.tail = FALSE) + BayesPower:::p_cor(min(r), location, n, lower.tail = TRUE) else if (r > location) BayesPower:::p_cor(r, location, n, lower.tail = FALSE) else BayesPower:::p_cor(r, location, n, lower.tail = TRUE),
                "greater" = BayesPower:::p_cor(r, location, n, lower.tail = FALSE),
                "less" = BayesPower:::p_cor(r, location, n, lower.tail = TRUE))
    return(x)
  }

  # limits of the integral and the parameter space
  bound <- switch(alternative, "greater" = c(a = h0, b = 1), "less" = c(a = -1, b = h0), "two.sided" = c(a = -1, b = 1))

  # normalization of prior under h1
  normalization <- if (alternative == "two.sided") {
    switch(prior_analysis, "d_beta" = 1, "beta" = 1, "Moment" = BayesPower:::pmom(bound[2] - location, tau = scale^2) - BayesPower:::pmom(bound[1] - location, tau = scale^2))
  } else {
    switch(prior_analysis, "d_beta" = BayesPower:::p_beta(bound[2], 1 / k, 1 / k, -1, 1) - BayesPower:::p_beta(bound[1], 1 / k, 1 / k, -1, 1), "beta" = BayesPower:::p_beta(bound[2], alpha, beta, -1, 1) - BayesPower:::p_beta(bound[1], alpha, beta, -1, 1), "Moment" = BayesPower:::pmom(bound[2] - location, tau = scale^2) - BayesPower:::pmom(bound[1] - location, tau = scale^2))
  }

  # integral
  int <- function(rho) {
    prob <- switch(alternative,
                   "two.sided" = if (length(r) == 2) BayesPower:::p_cor(max(r), rho, n, lower.tail = FALSE) + BayesPower:::p_cor(min(r), rho, n, lower.tail = TRUE) else if (r > location) BayesPower:::p_cor(r, rho, n, lower.tail = FALSE) else BayesPower:::p_cor(r, rho, n, lower.tail = TRUE),
                   "greater" = BayesPower:::p_cor(r, rho, n, lower.tail = FALSE),
                   "less" = BayesPower:::p_cor(r, rho, n, lower.tail = TRUE))

    prob * BayesPower:::r_prior(rho, k, location, scale, dff, prior_analysis, alpha, beta, min(bound), max(bound)) / normalization
  }

  #x <- stats::integrate(int, lower = bound[1], upper = bound[2], rel.tol = 1e-4)$value
  x <- stats::integrate(int,lower = bound[1],upper = bound[2], rel.tol = 1e-4)$abs.error
  return(x)

}

```

### Analysis settings

This section defines the prior specifications, sample-size settings, and critical values used in the numerical comparison. 

```{r, results='hide'}
r.cri          <- seq(0.01,.5,.01)
n              <- 50
k              <- 1
alpha          <- beta     <- 2
h0             <- location <- 0
scale          <- .01
dff            <- 1
prior_analysis <-  c("d_beta","beta","Moment")
```

### Calculating absolute error across different correlations

```{r}
abs_error_r    <- array(NA, dim=c(3,length(r.cri)))

for( ii in 1:length(prior_analysis)){
  for( i in 1:length(r.cri)){
    r <- c(-r.cri[i],r.cri[i]) # specifying which critical value is used for power calculation
    abs_error_r[ii,i] <- r_TPE_abs_error(r,n,k, alpha, beta,h0,alternative,location,scale,dff,prior_analysis[ii])
  }}
max(abs_error_r)
```




## $f$-test

The function in the backend for the calculation of power, `F_TPE`, is modified such that they return the absolute error from the numerical integration instead of the estimated value of the integral.


```{r, results='hide'}
##### Helper #####

F_TPE_abs_error<-function(f,q,m,dff,rscale,f_m,prior_analysis){
  if (length(f) == 0 || any(f == "no bound is found")) return(0)

  if (prior_analysis == "Point"){
    x = stats::pf(f,q,m-q,ncp =m*f_m^2,lower.tail = F)
    return(x)
  }
  # integral, the marginalized probability
  int  <- function(fsq){

    stats::pf(f,q,m-q,ncp =m*fsq,lower.tail = F)*BayesPower:::F_prior(fsq,q,dff,rscale,f_m,prior_analysis)
  }
  x = stats::integrate(int,lower = 0,upper = Inf)$abs.error

  return(x)
}


```

### Analysis settings

This section defines the prior specifications, sample-size settings, and critical values used in the numerical comparison. 

```{r, results='hide'}
f.cri          <- seq(0.1,5,.01)
k              <- 10
p              <- 2
q              <- k-p
n              <- 50
m              <- n-p
prior_analysis <- c("effectsize", "Moment")
dff            <- 3
rscale         <- .1
f_m            <- sqrt(.1)
```

### Calculating absolute error across different $f$-value

```{r}
abs_error_f    <- array(NA, dim=c(2,length(f.cri)))
for( ii in 1:length(prior_analysis)){
  for( i in 1:length(f.cri)){
    f <- f.cri[i] # specifying which critical value is used for power calculation
    abs_error_f[ii,i] <- F_TPE_abs_error(f,q,m,dff,rscale,f_m,prior_analysis[ii])

  }}
max(abs_error_f)
```






## One Proportion

The function in the backend for the calculation of power, `bin_TPE`, is modified such that they return the absolute error from the numerical integration instead of the estimated value of the integral.


```{r, results='hide'}
##### Helper #####

bin_TPE_abs_error<-function(x,n,h0,alpha,beta,location,scale,prior_analysis,alternative){
  if (length(x) == 0 || any(x == "bound cannot be found")) return(0)

  if (prior_analysis =="Point"){
    TPE = switch(alternative,
                 "two.sided" = {

                   switch(length(x)==2,
                          "1" ={stats::pbinom(min(x),n,location,lower.tail = T)+ stats::pbinom(max(x)-1,n,location,lower.tail = F)},
                          "0"=  {
                            switch(x/n>location,
                                   "1" = stats::pbinom(x-1,n,location,lower.tail = F),
                                   "0" = stats::pbinom(x,n,location,lower.tail = T))

                          })
                 },
                 "greater"  = {stats::pbinom(x-1,n,location,lower.tail = F)},
                 "less"  = {stats::pbinom(x,n,location,lower.tail = T)}
    )
    return(TPE)
  }

  # limits of the integral and the parameter space
  bound  <- switch(alternative,
                   "greater" = c(a = h0, b = 1),
                   "less" = c(a = 0, b = h0),
                   "two.sided" = c(a = 0, b = 1)
  )
  # normalization of the prior under h1
  normalization <- if (alternative == "two.sided") {
    switch(prior_analysis,
           "Moment"   = BayesPower:::pmom(bound[2]-location, tau=scale^2)-BayesPower:::pmom(bound[1]-location, tau=scale^2),
           "beta"     = 1)

  } else {
    switch(prior_analysis,
           "Moment"   = BayesPower:::pmom(bound[2]-location, tau=scale^2)-BayesPower:::pmom(bound[1]-location, tau=scale^2),
           "beta"     = stats::pbeta(bound[2],alpha,beta)-stats::pbeta(bound[1],alpha,beta))
  }
  # integral, the probability is marginalized with respect to the prior, depending on the alternative hypothesis,h0 and the length of the critical values.
  int <- function(prop) {
    pro <- switch(alternative,
                  "two.sided" = {
                    if (length(x) == 2) {
                      stats::pbinom(min(x), n, prop, lower.tail = TRUE) +
                        stats::pbinom(max(x) - 1, n, prop, lower.tail = FALSE)
                    } else {
                      mapply(function(x_i, n_i, p_i) {
                        if (x_i / n_i > location) {
                          stats::pbinom(x_i - 1, n_i, p_i, lower.tail = FALSE)
                        } else {
                          stats::pbinom(x_i, n_i, p_i, lower.tail = TRUE)
                        }
                      }, x, n, prop)
                    }
                  },
                  "greater" = stats::pbinom(x - 1, n, prop, lower.tail = FALSE),
                  "less" = stats::pbinom(x, n, prop, lower.tail = TRUE)
    )

    pro * BayesPower:::bin_prior(prop, alpha, beta, location, scale, prior_analysis) / normalization
  }

  TPE = stats::integrate(int,lower = bound[1],upper = bound[2], rel.tol = 1e-5)$abs.error

  return(TPE)

}


```

### Analysis settings

This section defines the prior specifications, sample-size settings, and critical values used in the numerical comparison. 

```{r, results='hide'}
n               <- 50
h0              <- location  <-.5
alpha           <- beta      <- 1
scale           <- .1
alternative     <- "two.sided"
prior_analysis  <- c("beta","Moment")
# specifying which critical values in terms of number of success are used for power calculation
lower          <-  24:0   
upper          <-  26:50  

```

### Calculating absolute error across different number of success

```{r}
abs_error_bin   <- array(NA, dim=c(2,25))
for( ii in 1:length(prior_analysis)){
  for( i in 1:(25)){
    x <- c(lower[i],upper[i]) # specifying which critical value is used for power calculation
    abs_error_bin[ii,i] <- bin_TPE_abs_error(x,n,h0,alpha,beta,location,scale,prior_analysis[ii],alternative)


  }}
max(abs_error_bin)
```

## Maximum error across all tests
The maximum absolute error across all tests is obtained by:

```{r}
max(abs_error_t1,abs_error_t2,abs_error_r,abs_error_f,abs_error_t1,abs_error_bin)

```
The results when excluding the one for $f$-test with effect size analysis prior
```{r}
max(abs_error_t1,abs_error_t2,abs_error_r,abs_error_f[2,],abs_error_t1,abs_error_bin)
```
